---
title: "Build a hierarchical reserving model in R with hirem"
author: "Jonas Crevecoeur"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hirem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{css, echo = FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

<style>
body {
text-align: justify}
</style>

```{r require}
require(hirem) 
require(tidyr)
require(magrittr)
require(dplyr)
require(tidyverse)
```

## Reserving data set included in hirem
The `hirem` package includes a data set, which is simualted based on a real insurance portfolio.

```{r data}
data('reserving_data')
head(reserving_data)
```

In this data set the development of a claim in a development year is recorded in three variables

* close: One when the claim settles in the curent development year, zero otherwise
* payment: One when there is a payment in the current development year, zero otherwise
* size: Size of the payment in the current development year

The data set `reserving_data` is not censored. We artificially create the censoring that one would typically find in reserving data

```{r}
upper_triangle <- reserving_data %>% filter(calendar_year <= 6)
lower_triangle <- reserving_data %>% filter(calendar_year > 6)
```

## Building the hierarchical reserving model

The `hirem` package starts with an empty model and constructs the hierarchical model by sequentially adding new layers. Each layer represents an event recorded in the data set.

The function `hirem` start a new hierarchical reserving model. This function has a single argument `data` to which we pass the data on which we want to train the model.

```{r, eval = FALSE}
model <- hirem(upper_triangle)
```

Two types of layers are currently implemented in the `hirem` package

* layer_glm: Adds a layer estimated using generalized linear models
* layer_gbm: Adds a layer estimated using gradient boosting models

We construct a hierarchical reserving model using `layer_glm`. See the documentation for a detailed description on the use of `?layer_gbm`.

`layer_glm` requires 2 parameters:

* name: The name of the layer. This name has to match the column name used of the covariate in the input data set.
* family: The family object to pass to the `glm` routine.

The layer `size` is zero when there is no payment. To include this relation in the hierarchical reserving model, we add the optional parameter `filter`. `filter` is a function that returns a vector with values `TRUE`/`FALSE`. Records for which `filter` evaluates to `FALSE` are not included when fitting the layer and are set to zero in the simulation.

```{r}
model <- hirem(upper_triangle) %>%
  layer_glm('close', binomial(link = cloglog)) %>%
  layer_glm('payment', binomial(link = logit)) %>%
  layer_glm('size', Gamma(link = log),
            filter = function(data){data$payment == 1})
```

Call `fit` to calibrate these generalized linear models. We pass to this function the formula describing the regresion model for each component.
```{r}
model <- fit(model,
             close = 'close ~ factor(development_year) + factor(X1) + factor(X2)',
             payment = 'payment ~ close + factor(development_year) + factor(X1) + factor(X2)',
              size = 'size ~ close + factor(development_year) + factor(X1) + factor(X2)')
```

## Simulate the future reserve

We have now defined and trained a hierarchical reserving model. This model defines the evolution of the events `close`, `payment` and `size` in a claims lifetime. Besides, these stochastic covariates, there are other covariates that have a deterministic evolution over time (e.g. `development_year`). 

Before, we can simulate paths for the future development of claims, we register an `updater` to `update` these deterministic covariates.

```{r}
update <- function(data) {
  data %>%
    dplyr::mutate(development_year = development_year + 1,
                  calendar_year = calendar_year + 1)
}

model <- register_updater(model, update)
```

See the documentation (?register_updater) for more information. By default the data is updated at the end of each cycle (development_year), but it is also possible to add updates after a specific layer. These updates can keep variables such as the total amount paid up to date during the simulation. 

Call `simulate` to simulate future paths for reported claims. This function has 4 arguments:

* obj: The hierarchical model from which we want to simulate
* nsim: The number of simulations
* filter: A function removing all claims for which we don't want to simualte the next development year. 
* data: The last observed status for the claims for which we want to simualte future paths

In our example, we simulate claims until they settle or until development year 6

```{r}
simul <- simulate(model,
                  nsim = 5,
                  filter = function(data){dplyr::filter(data,
                                                       development_year <= 6,
                                                       close == 0)},
                  data = reserving_data %>% dplyr::filter(calendar_year == 6))
```

This simulation has the same structure as the input data

```{r}
head(simul)
```

The extra columns `simulation` identifies the different simulations. 

Comparing the predicted and actual reserve

```{r}
rbns_estimate <- simul %>%
  dplyr::group_by(simulation) %>%
  dplyr::summarise(rbns = sum(size))

rbns_actual <- reserving_data %>%
  dplyr::filter(calendar_year > 6) %>%
  dplyr::summarise(rbns = sum(size))

rbns_estimate
rbns_actual
```

Predicting the lower half of the runoff triangle
```{r}

lower_triangle_predicted <- simul %>%
  dplyr::group_by(reporting_year, development_year) %>%
  dplyr::summarise(total_size = sum(size) / max(simulation)) %>%
  dplyr::arrange(development_year) %>%
  tidyr::pivot_wider(values_from = total_size, names_from = development_year) %>%
  dplyr::arrange(reporting_year)

lower_triangle_actual <- reserving_data %>%
  dplyr::filter(calendar_year > 6) %>%
  dplyr::group_by(reporting_year, development_year) %>%
  dplyr::summarise(total_size = sum(size)) %>%
  dplyr::arrange(development_year) %>%
  tidyr::pivot_wider(values_from = total_size, names_from = development_year) %>%
  dplyr::arrange(reporting_year)

lower_triangle_actual
lower_triangle_predicted

```

